Shaping the Perception of a Robot by Cognitive Priming

### 2016-09-02 main.tex
{shallow/deep -> machine-like/human-like} + {A0/A1 -> Astart/Aend}

\newcommand{\anti}{{$\mathcal{A}_0$\xspace}}
\newcommand{\antf}{{$\mathcal{A}_1$\xspace}}
\newcommand{\deltaant}{{ $\Delta_{\mathcal{A}_0,\mathcal{A}_1}$\xspace}}

@@

All the above studies show that the cognitive and/or affective priming might alter the behaviour and/or the perception of the subjects about a target stimulus. Some of the studies found some complex interactions between cognitive and affective priming (\cite{yi1990cognitive} and \cite{storbeck2008affective} ). In this contribution, we were only interested in the cognitive priming aspect during a Human Robot Interaction situation, as \cite{baldwin1990priming} advocated that the cognitive priming may activate a relationship schema between the priming stimulus and the target stimulus. Moreover, in the context of advertisement, a number of studies suggest that the ad context (cognitive priming) can influence the audience's perception of the ad, and thus, its effectiveness (e.g., \cite{singh1987arousal};\cite{soldow1981response})  Our primary hypothesis was that the way the robot is given commands might alter the way the subjects perceive its actions. We propose to use a neutral mood (no affective priming) and two cognitive priming conditions (Deep and Shallow), in order to remove the ``mood'' or  ``affect'' bias of the priming. The ``shallow'' priming consisted of giving commands to the robot which are more ``machine-like''. On the other hand, the ``deep'' priming had the commands to robot in a more ``human-like'' manner (Manipulation-check hypothesis). We also kept the time of priming to be the same in the deep and shallow conditions, in order to remove the time of priming effect on the subjects' perceptions about the robot.

@@

This study is built as an eye-tracking experiment (using a stationary
eye-tracker SMI RED 250) where participants watch short videos of an agent --
either a robot (Aldebaran Robotics' Nao) or a human -- performing simple
tasks (picking an object or pointing toward a sound source) after being primed.
The priming is designed to elicit either shallow cognitive ascriptions onto the
robot (leading to a \emph{machine-like} situation) or deeper cognitive
ascriptions (leaning toward a \emph{human-like} situation).

@@

The study follows a 2$\times$2 design, summarised in figure~\ref{table:design}.
Our two independent variables are the level of cognitive priming (\emph{shallow}
\vs \emph{deep}) and the nature of the agent appearing in the stimulus (a robot
or a human).

@@

    \tikzmark{left}
    \begin{tabular}{l|p{4cm}|p{4cm}}
        & \tikzmark{top1} Shallow cognitive priming & Deep cognitive priming \tikzmark{top2}  \\
        \hline
        \tikzmark{left1} Robot & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''} \\
        \hline
        \tikzmark{left2} Human & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\ 
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''}\tikzmark{bottom} \\
        \end{tabular}

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive priming} \vs \emph{deep cognitive priming} is between subjects.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

The audio (and audio only) of these four videos was then edited to create two
sets of stimuli: one for the \emph{shallow cognitive priming} condition, one for
the \emph{deep cognitive priming} condition. The audio editing consisted of 
inserting different commands to initiate the tasks (reported in
table~\ref{table:design}): in the \emph{Picking} task for instance, the object
that the agent had to pick was described either as a \emph{``brown toy''} or as
\emph{``your favourite toy''}. In the former case, the object was referred through
visible, objective features. They do not presuppose as deep cognitive capabilities
as in the latter case where the priming is expected to elicit the assumption
that the robot has tastes and preferences such as it can choose a toy as its favourite.

This simple verbal prompt (\emph{``brown toy''} \vs \emph{``your favourite
toy''}) set the expectations regarding the cognitive capabilities of the robot
(\emph{shallow} \vs \emph{deep}) and is expected to lead to different
perceptions of the robot behaviour (\emph{machine-like} \vs \emph{human-like}). We
reinforced this effect in the \emph{Pointing} task: not only the verbal prompt
(\emph{``point at the noise''} \vs \emph{``point at the crying baby''}) but also
the background sound played by the speakers during the task was altered: either a
repetitive beep (\emph{machine-like} context) or the sound of a crying baby
(\emph{human-like} context).

@@

As stated previously, the study followed a 2$\times$2 design
(table~\ref{table:design}): the \emph{shallow} \vs \emph{deep cognitive priming}
condition was between-subjects (30 participants performed the \emph{deep cognitive}
tasks while 26 were in the \emph{shallow cognitive} condition); the \emph{robot}
\vs \emph{human} condition was within-subject (the order of stimulus presentation
was counter-balanced over the participants). The study lasted in average 15
minutes per participant.

@@

{\Large
\[
    \delta_{\mathcal{H}, \mathcal{R}}^{\text{gaze}} =
    \frac{\sum\limits_{{\text{{\sc aoi}s}}}\left |
    \frac{\text{Time}_{\text{Human}}^{\text{\sc
aoi}}}{\text{Area}_{\text{Human}}^{\text{\sc aoi}}} -
\frac{\text{Time}_{\text{Robot}}^{\text{\sc
aoi}}}{\text{Area}_{\text{Robot}}^{\text{\sc aoi}}} \right |}{N}
\]
}

@@

Based on the experiments conducted on 56 participants in the human and robot
interactions across \emph{shallow} and \emph{deep cognitive priming} conditions,
we have the following results.\footnote{The detailed statistics and code used
    for obtaining the results are available here :
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{shallow}=TBD (SD=TBD)$
\vs $M_{deep}=TBD (SD=TDB)$) in this
manipulation check ($F(1,54) = 29.0, p < .001$) which validates our manipulation
and supports \h{1}: the participants that watched stimuli with a \emph{shallow cognitive
priming} identified the task as more robot-like and participants watching
the videos with a deep cognitive priming perceived the tasks as more human-like.

@@

\subsection{Gaze Behaviours in Response to Shallow \vs Deep Cognitive Priming}

We consider the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf hands}, {\sf
torso} and {\sf legs} (the left and right arms, hands and legs are grouped
together, other AOIs not belonging to the robot are left out). Based on ANOVA results for the robot in both the shallow and deep
cognitive priming conditions, the fixations on the head in the deep condition
were found to be significantly higher than that in the shallow condition
($F(1,54) = 4.2, p < .05$\fixme{put the exact p value here}). For all other AOIs groups, the fixations across
shallow and deep conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.6\columnwidth]{GazeHighLow}\label{GazeHighLow}
    \caption{Gaze fixations on the different areas on interest of the robot, for shallow \vs
    deep cognitive priming conditions. TBD: what do the error bar represent?}
    \label{h3}
\end{figure}

@@

This constitutes the main novel result of this study: priming an observer with a deeper cognitive
context for the same interaction situation suffices to lead to significantly longer
fixations on the robot's head (hypothesis \h{3}).

Note that we only analysed these possible correlations for robot condition: in
the human condition, the presence of the human is intrinsically a source of deep
cognitive expectations, hiding the effects of our manipulation on a shallow \vs deep
cognitive priming.

@@

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for tasks with a deeper cognitive
context than that for the shallow cognitive tasks ($\Delta_{shallow}=TBD (SD=TBD)$
\vs $\Delta_{deep}=TBD (SD=TDB)$, $F(1,54) = 6.54, p < .05$\fixme{put the exact
p value here}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with a deeper cognitive
expectations report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

%\begin{figure}
%    \centering
%    \includegraphics[width=3.3in]{H4}\label{ICAtoAAPImprovement}
%    \caption{Difference between \anti{} and \antf{} in the shallow and deep
%    cognitive priming conditions.}
%    \label{h4}
%\end{figure}

This result supports \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots \emph{beyond the specific Nao robot
shown in the stimuli}.  It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{deep} cognitive condition to ``robots'' in general (we can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment).

@@

The significant impact of the deep \vs shallow commands implication on the
answers to the question \emph{Were the tasks more robot-like or human-like?}
shows that our manipulation was well recognized, and that a shallow cognitive
context was indeed perceived to elicit a ``robot-like'' situation, while a deeper
cognitive priming was perceived as a ``human-like'' situation. This supports our
ifirst hypothesis \h{1}: even indirect hints at the cognitive capabilities of a robot are
sufficient to elicit significantly increased attribution of human-likeness.

@@

The third hypothesis \h{3} proposes that head fixations are related to the
cognitive expectations, which is clearly supported by our data: after deeper cognitive
priming, one tends to spend more time looking at the head of the interacting
agent than after shallow cognitive priming. We also hypothesised that, in turn,
head fixations could act as a \emph{proxy to anthropomorphic projections}: as
underlined in the previous section, our data also support this.

Finally, the hypothesis \h{4} looks at the dynamic of anthropomorphic
projections, with the hypothesis that watching an interaction with a particular
robot after a deeper cognitive priming may impact \emph{globally} the
anthropomorphic perceptions of robots. This is supported by our study.

@@

Finally, the situations needed to support two possible kind of cognitive priming
(shallow \vs deeper cognitive priming), and triggering one or the other of these
contexts needed to be done \emph{without changing} the visual part of the
stimulus.

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Pointing} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{shallow
cognitive priming} condition) or to \emph{``Dance with the music''}
(\emph{deeper cognitive priming} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as we
found out that the nature of the task (participants had to watch body movements)
introduced an experimental bias: because dance is a form of whole-body
expression, gazes where evenly distributed on the body of the agents, masking
possible cognition-related gaze patterns.

@@

This leads to a second-order effect that we also evidence here: depending on the
cognitive priming, the tendency to attribute human-like characteristics to robot
evolves in significantly different ways. After observing a robot in a
context that presupposes deeper cognitive capabilities, people increase
their general tendency to anthropomorphise robots compared to a shallow
cognitive priming, even though the robot appearance and visual behaviour is the
exact same.

>>>>

\newcommand{\anti}{{$\mathcal{A}_{start}$\xspace}}
\newcommand{\antf}{{$\mathcal{A}_{end}$\xspace}}
\newcommand{\deltaant}{{$\Delta_{\mathcal{A}}$\xspace}}

@@

All the above studies show that the cognitive and/or affective priming might alter the behaviour and/or the perception of the subjects about a target stimulus. Some of the studies found some complex interactions between cognitive and affective priming (\cite{yi1990cognitive} and \cite{storbeck2008affective} ). In this contribution, we were only interested in the cognitive priming aspect during a Human Robot Interaction situation, as \cite{baldwin1990priming} advocated that the cognitive priming may activate a relationship schema between the priming stimulus and the target stimulus. Moreover, in the context of advertisement, a number of studies suggest that the ad context (cognitive priming) can influence the audience's perception of the ad, and thus, its effectiveness (e.g., \cite{singh1987arousal};\cite{soldow1981response})  Our primary hypothesis was that the way the robot is given commands might alter the way the subjects perceive its actions. We propose to use a neutral mood (no affective priming) and two cognitive priming conditions (Human-like and Machine-like), in order to remove the ``mood'' or  ``affect'' bias of the priming. The ``machine-like'' priming consisted of giving commands to the robot which are more ``machine-like''. On the other hand, the ``human-like'' priming had the commands to robot in a more ``human-like'' manner (Manipulation-check hypothesis). We also kept the time of priming to be the same in the human-like and machine-like conditions, in order to remove the time of priming effect on the subjects' perceptions about the robot.

@@

This study is built as an eye-tracking experiment (using a stationary
eye-tracker SMI RED 250) where participants watch short videos of an agent --
either a robot (Aldebaran Robotics' Nao) or a human -- performing simple tasks
(picking an object or pointing toward a sound source) after being primed.  The
priming is designed to elicit either low cognitive ascriptions onto the robot
(leading to what we call hereafter a \emph{machine-like} interaction situation)
or richer cognitive ascriptions (leaning toward what we call a \emph{human-like}
interaction situation).

@@

The study follows a 2$\times$2 design, summarised in figure~\ref{table:design}.
Our two independent variables are the level of cognitive priming (\emph{machine-like}
\vs \emph{human-like}) and the nature of the agent appearing in the stimulus (a robot
or a human).

@@

    \tikzmark{left}
    \begin{tabular}{l|p{4cm}|p{4cm}}
        & \tikzmark{top1} Machine-like cognitive priming & Human-like cognitive priming \tikzmark{top2}  \\
        \hline
        \tikzmark{left1} Robot & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''} \\
        \hline
        \tikzmark{left2} Human & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\ 
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''}\tikzmark{bottom} \\
        \end{tabular}

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{machine-like
        cognitive priming} \vs \emph{human-like cognitive priming} is between subjects.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

The audio (and audio only) of these four videos was then edited to create two
sets of stimuli: one for the \emph{machine-like cognitive priming} condition, one for
the \emph{human-like cognitive priming} condition. The audio editing consisted of 
inserting different commands to initiate the tasks (reported in
table~\ref{table:design}): in the \emph{Picking} task for instance, the object
that the agent had to pick was described either as a \emph{``brown toy''} or as
\emph{``your favourite toy''}. In the former case, the object was referred through
visible, objective features. They do not presuppose as complex cognitive capabilities
as in the latter case where the priming is expected to elicit the assumption
that the robot has tastes and preferences such as it can choose a toy as its favourite.

This simple verbal prompt (\emph{``brown toy''} \vs \emph{``your favourite
toy''}) creates a set of expectations regarding the cognitive capabilities of
the robot and leads to a different perceptions of the interaction situation (the
task performed by the robot is seen by the participant as more
\emph{machine-like} or more \emph{human-like}). We reinforced this effect in the
\emph{Pointing} task: not only the verbal prompt (\emph{``point at the noise''}
\vs \emph{``point at the crying baby''}) but also the background sound played by
the speakers during the task was altered: either a repetitive beep
(\emph{machine-like} situation) or the sound of a crying baby (\emph{human-like}
situation).

@@

As stated previously, the study followed a 2$\times$2 design
(table~\ref{table:design}): the \emph{machine-like} \vs \emph{human-like cognitive priming}
condition was between-subjects (30 participants performed the \emph{human-like cognitive}
tasks while 26 were in the \emph{machine-like cognitive} condition); the \emph{robot}
\vs \emph{human} condition was within-subject (the order of stimulus presentation
was counter-balanced over the participants). The study lasted in average 15
minutes per participant.

@@

{\Large
\[
    \delta_{\mathcal{H}, \mathcal{R}}^{\text{gaze}} =
    \frac{\sum\limits_{{\text{{\sc aoi}s}}}\left |
    \frac{\text{Duration}_{\text{Human}}^{\text{\sc
aoi}}}{\text{Area}_{\text{Human}}^{\text{\sc aoi}}} -
\frac{\text{Duration}_{\text{Robot}}^{\text{\sc
aoi}}}{\text{Area}_{\text{Robot}}^{\text{\sc aoi}}} \right |}{N}
\]
}

@@

Based on the experiments conducted on 56 participants in the human and robot
interactions across \emph{machine-like} and \emph{human-like cognitive priming} conditions,
we have the following results.\footnote{The detailed statistics and code used
    for obtaining the results are available here :
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{\text{machine-like}}=TBD (SD=TBD)$
\vs $M_{\text{human-like}}=TBD (SD=TDB)$) in this
manipulation check ($F(1,54) = 29.0, p < .001$) which validates our manipulation
and supports \h{1}: the participants that watched stimuli with a \emph{machine-like cognitive
priming} identified the task as more robot-like and participants watching
the videos with a human-like cognitive priming perceived the tasks as more human-like.

@@

\subsection{Gaze Behaviours in Response to Machine-like \vs Human-like Cognitive Priming}

We consider the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf hands}, {\sf
torso} and {\sf legs} (the left and right arms, hands and legs are grouped
together, other AOIs not belonging to the robot are left out). Based on ANOVA results for the robot in both the machine-like and human-like
cognitive priming conditions, the fixations on the head in the human-like condition
were found to be significantly higher than that in the machine-like condition
($F(1,54) = 4.2, p < .05$\fixme{put the exact p value here}). For all other AOIs groups, the fixations across
machine-like and human-like conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.6\columnwidth]{GazeHighLow}\label{GazeHighLow}
    \caption{Gaze fixations on the different areas on interest of the robot, for machine-like \vs
    human-like cognitive priming conditions. TBD: what do the error bar represent?}
    \label{h3}
\end{figure}

@@

This constitutes the main novel result of this study: priming an observer with
the expectation of a robot endowed with richer cognitive abilities leads to significantly longer
fixations on the robot's head, even though the actual behaviour of the robot is
the same (hypothesis \h{3}).

Note that we only analysed these possible correlations for robot condition: in
the human condition, the presence of the human is intrinsically a source of rich
cognitive expectations, masking the effects of our manipulation on a
machine-like \vs human-like cognitive priming.

@@

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for the participants who watched the
human-like tasks ($\Delta_{\text{machine-like}}=TBD (SD=TBD)$
\vs $\Delta_{\text{human-like}}=TBD (SD=TDB)$, $F(1,54) = 6.54, p < .05$\fixme{put the exact
p value here}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with a richer cognitive
expectations regarding the robot report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

%\begin{figure}
%    \centering
%    \includegraphics[width=3.3in]{H4}\label{ICAtoAAPImprovement}
%    \caption{Difference between \anti{} and \antf{} in the machine-like and
%    human-like
%    cognitive priming conditions.}
%    \label{h4}
%\end{figure}

This result supports \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots \emph{beyond the specific Nao robot
shown in the stimuli}.  It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{human-like} cognitive condition to ``robots'' in general (we can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment).

@@

The significant impact of the machine-like \vs human-like conditions on the
answers to the question \emph{Were the tasks more ``a robot kind of task'' or
``a human kind of task''?} shows that our manipulation was effective: framing
the interaction situation to induce a machine-like or a human-like task context
was successful.

situation priming the participants with , and that a machine-like cognitive
context was indeed perceived to elicit a ``robot-like'' situation, while a
human-like
cognitive priming was perceived as a ``human-like'' situation. This supports our
first hypothesis \h{1}: even indirect hints at the cognitive capabilities of a robot are
sufficient to elicit significantly increased attribution of human-likeness.

@@

The third hypothesis \h{3} proposes that head fixations are related to the
cognitive expectations, which is clearly supported by our data: after human-like cognitive
priming, one tends to spend more time looking at the head of the interacting
agent than after machine-like cognitive priming. We also hypothesised that, in turn,
head fixations could act as a \emph{proxy to anthropomorphic projections}: as
underlined in the previous section, our data also support this.

Finally, the hypothesis \h{4} looks at the dynamic of anthropomorphic
projections, with the hypothesis that watching an interaction with a particular
robot after a human-like cognitive priming may impact \emph{globally} the
anthropomorphic perceptions of robots. This is supported by our study.

@@

Finally, the situations needed to support two possible kind of cognitive priming
(machine-like \vs human-like cognitive priming), and triggering one or the other of these
contexts needed to be done \emph{without changing} the visual part of the
stimulus.

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Pointing} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{machine-like
cognitive priming} condition) or to \emph{``Dance with the music''}
(\emph{human-like cognitive priming} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as we
found out that the nature of the task (participants had to watch body movements)
introduced an experimental bias: because dance is a form of whole-body
expression, gazes where evenly distributed on the body of the agents, masking
possible cognition-related gaze patterns.

@@

This leads to a second-order effect that we also evidence here: depending on the
cognitive priming, the tendency to attribute human-like characteristics to robot
evolves in significantly different ways. After observing a robot in a
context that presupposes richer cognitive capabilities, people increase
their general tendency to anthropomorphise robots compared to a machine-like
cognitive priming, even though the robot appearance and visual behaviour is the
exact same.
